---
title: "US & Global Energy Forecast Data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# importing necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
```


```{r data import, include = FALSE}
# total energy production data
tot_energy_production <- read.csv("data/Total_Energy_Production.csv")

# total energy use data
tot_energy_use <- read.csv("data/Total_Energy_Use.csv")

# total energy imports data
tot_energy_imports <- read.csv("data/Total_Energy_Imports.csv")

# total energy exports data
tot_energy_exports <- read.csv("data/Total_Energy_Exports.csv")

# total energy real prices data
tot_energy_real_prices <- read.csv("data/Total_Energy_Real_Prices.csv")

# CO2 emissions by Fuel type
emissions_by_fuel_type <- read.csv("data/Carbon_Dioxide_Total_by_Fuel.csv")

# Renewable Energy generating capacity
renewable_capacity <- read.csv("data/Renewable_Energy_Generating_Capacity.csv")

# renewable energy capacity additions
renewable_additions <- read.csv("data/Renewable_Energy_Capacity_Additions.csv")

# energy use oecd vs non-oecd
global_energy_use <- read.csv("data/energy_use_OECD_non_OECD.csv")

# emissions by OECD
emissions_from_OECD <- read.csv("data/Carbon_dioxide_emissions_OECD.csv")

# emissions non OECD
emissions_from_non_OECD <- read.csv("data/Carbon_dioxide_emissions_non_OECD.csv")
```

```{r data transformation, include = FALSE}
#transforming energy production data - LETS GOOOO
new_energy_production <- tot_energy_production %>%
  select(year = Year, 
         liquidfuels = Crude.Oil.and.Lease.Condensate.quads,
         naturalgas = Natural.Gas.Plant.Liquids.quads,
         coal = Coal.quads,
         nuclear = Nuclear.quads,
         hydropower = Hydropower.quads,
         biomass = Biomass.quads,
         other_renewable_energy = Other.Renewable.Energy.quads,
         other = Other.quads,
         total = Total.quads) %>%
  pivot_longer(c("liquidfuels", "naturalgas", "coal", "nuclear", "hydropower", 
                 "biomass", "other_renewable_energy", "other", "total"), 
               names_to = "fueltype", values_to = "quads") %>%
    drop_na() # removing any NA values to avoid error message


# Checking new df
new_energy_production

# transforming energy consumption data
new_energy_use <- tot_energy_use %>%
  select(year = Year, 
         liquidfuels = Liquid.Fuels.quads,
         naturalgas = Natural.Gas.quads,
         coal = Coal.quads,
         nuclear = Nuclear.quads,
         hydropower = Hydropower.quads,
         biomass = Biomass.quads,
         other_renewable_energy = Other.Renewable.Energy.quads,
         other = Other.quads,
         total = Total.quads) %>%
  pivot_longer(c("liquidfuels", "naturalgas", "coal", "nuclear", "hydropower", 
                 "biomass", "other_renewable_energy", "other", "total"), 
               names_to = "fueltype", 
               values_to = "quads") %>%
    drop_na()

# checking new df
new_energy_use

# transforming import data
new_energy_imports <- tot_energy_imports %>%
  select(year = Year, 
         crudeoil = Crude.Oil.quads,
         liquidfuels = Liquid.Fuels.quads,
         naturalgas = Natural.Gas.quads,
         other = Other.quads,
         total = Total.quads) %>%
  pivot_longer(c("crudeoil", "liquidfuels", "naturalgas", "other", "total"), 
               names_to = "fueltype", 
               values_to = "quads") %>%
    drop_na()

# check new imports df
new_energy_imports

# transforming export data
new_energy_exports <- tot_energy_exports %>%
  select(year = Year,
         petroleum = Petroleum.quads,
         naturalgas = Natural.Gas.quads,
         coal = Coal.quads,
         total = Total.quads) %>%
  pivot_longer(c("petroleum", "naturalgas", "coal", "total"), 
               names_to = "fueltype", 
               values_to = "quads") %>%
    drop_na()

# checking new exports df
new_energy_exports

# transforming real prices forecast
energy_real_prices <- tot_energy_real_prices %>%
  select(year = Year, 
         brent = Brent..2020...b..2020...b,
         wti = West.Texas.Intermediate..2020...b..2020...b,
         gas_henry_hub = Gas.Price.at.Henry.Hub..2020...MMBtu..2020...MMBtu,
         coal_minemouth_ton = Coal.at.Minemouth..2020...ton..2020...ton,
         coal_minemouth = Coal.Minemouth..2020...MMBtu..2020...MMBtu,
         deliver_coal = Delivered.Coal..2020...MMBtu..2020...MMBtu,
         electricity = Electricity..2020.cents.kWh..2020.cents.kWh) %>%
  pivot_longer(c("brent", "wti", "gas_henry_hub", "coal_minemouth", 
                 "coal_minemouth_ton", "deliver_coal","electricity"),
               names_to = "index", 
               values_to = "price") %>%
    drop_na()

# checking energy real prices df
energy_real_prices

# transform emissions data
emissions <- emissions_by_fuel_type %>%
  select(year = Year, 
         petroleum = Petroleum.MMmt.CO2,
         naturalgas = Natural.Gas.MMmt.CO2,
         coal = Coal.MMmt.CO2,
         other = Other.MMmt.CO2,
         total = Total.MMmt.CO2) %>%
  pivot_longer(c("petroleum", "naturalgas", "coal", "other", "total"), 
               names_to = "fueltype", 
               values_to = "mmotco2") %>%
    drop_na()

# checking emissions df
emissions

# transform renewable energy capacity
capacity <- renewable_capacity %>%
  select(year = Year, 
         hydropower = Hydropower.GW,
         geothermal = Geothermal.GW,
         mun_waste = Municipal.Waste.GW, 
         biomass = Wood.and.Other.Biomass.GW,
         solar_thermal = Solar.Thermal.GW,
         solar_pv = Solar.Photovoltaic.GW,
         wind = Wind.GW,
         offshore = Offshore.Wind.GW,
         total = Total.GW) %>%
  pivot_longer(c("hydropower", "geothermal", "mun_waste", "biomass", 
                 "solar_thermal", "solar_pv", "wind","offshore", "total"),
               names_to = "type", 
               values_to = "gw") %>%
    drop_na()

# check capacity df
capacity

# transform capacity additions 
cap_adds <- renewable_additions %>%
  select(year = Year, 
         hydropower = Hydropower.GW,
         geothermal = Geothermal.GW,
         mun_waste = Municipal.Waste.GW, 
         biomass = Wood.and.Other.Biomass.GW,
         solar_thermal = Solar.Thermal.GW,
         solar_pv = Solar.Photovoltaic.GW,
         wind = Wind.GW,
         offshore = Offshore.Wind.GW,
         total = Total.GW) %>%
  pivot_longer(c("hydropower", "geothermal", "mun_waste", "biomass", 
                 "solar_thermal", "solar_pv", "wind","offshore", "total"),
               names_to = "type", 
               values_to = "gw")%>%
    drop_na()

# check capacity additions df
cap_adds

# transform data for global energy use
energy_use_global <- global_energy_use %>%
  select(year = Year, 
         oecd_total = OECD..Total.OECD.quad.Btu,
         non_oecd_total = Non.OECD..Total.Non.OECD.quad.Btu) %>%
  pivot_longer(c("oecd_total", "non_oecd_total"),
               names_to = "region", 
               values_to = "quadbtu") %>%
    drop_na()

# check energy global use data frame
energy_use_global

# transform oecd emissions
oecd_emissions <- emissions_from_OECD %>%
  select(year = Year, 
         Americas = OECD.Americas.MMmt,
         Europe = OECD.Europe.MMmt,
         Asia = OECD.Asia.MMmt,
         total = Total.OECD.MMmt) %>%
  pivot_longer(c("Americas", "Europe", "Asia", "total"),
               names_to = "region",
               values_to = "MMmt") %>%
    drop_na()

# check oecd emissions df
oecd_emissions

# transform non oecd emissions
non_oecd_emissions <- emissions_from_non_OECD %>%
  select(year = Year, 
         eur = Non.OECD.Europe.and.Eurasia.MMmt,
         asia = Non.OECD.Asia.MMmt,
         middleeast = Middle.East.MMmt,
         africa = Africa.MMmt,
         americas = Non.OECD.Americas.MMmt,
         total = Total.Non.OECD.MMmt) %>%
  pivot_longer(c("eur", "asia", "middleeast", "africa", 
                 "americas",
                 "total"),
               names_to = "region",
               values_to = "MMmt")%>%
    drop_na()

# check oecd emissions df
non_oecd_emissions

```

```{r export import deep data transformation, include = FALSE}
# here, we want to put import/export data together in one df
exports <- new_energy_exports %>%
  mutate(direction = "Exports")

# show data frame
exports

# here we are doing the same but for import
imports <- new_energy_imports %>%
  mutate(direction = "Imports")

# show imports data frame now
imports

# joining the two data sets
export_import <- rbind(exports, imports)

# show joined data frame 
export_import

```

```{r renewables deep data transformation, include=FALSE}
# here we want to try and put capacity and capacity addition into one df
# adding established value to all rows
capacity_est <- capacity %>%
  mutate(status = "Established")

# show our df
capacity_est

# add addition value to all rows
capacity_addition <- cap_adds %>%
  mutate(status = "Addition")

# show our df
capacity_addition

# bind the two yaya
capacity_overview <- rbind(capacity_est, 
                           capacity_addition)

# show our binded df
capacity_overview

```


```{r US consumption/producton deep data transformation, include=FALSE}
# here we want to combine like we have before data sets
# for these dfs we will do it by if they are consumption or production values

consumption <- new_energy_use %>%
  mutate(result="Consumed")

production <- new_energy_production %>%
  mutate(result="Produced")

# stack our dfs on top of eachother 
usenergy <- rbind(consumption, production)
usenergy
```

```{r OECD and Non OECD deep data transformation, include = FALSE}
# our goal is to bind dfs for OECD and non OECD emissions and label 
# each respective row by its economic type

# first, OECD
oecd <- oecd_emissions %>%
  mutate(econ_type = "OECD")

# now, non-oecd
non_oecd <- non_oecd_emissions %>%
  mutate(econ_type = "Non-OECD")

# binding both dfs together 
world_emissions <- rbind(oecd, non_oecd)

```



#### **Energy**

##### The purpose of this report is to use publicly available data from the Energy
##### Information Administration (EIA) in order to think about the following questions:
1. Within a U.S. context, what do we expect energy to look like up until 2050?
2. Within a U.S. context, where is Renewable Energy headed?
3. Within a global context, what will energy and emissions look like thinking 
forward to 2050?

##### Before moving forward, it should be acknowledged that the data analyzed in this
##### report falls under the EIA 2020 Reference Case. This reference case maintains
##### very minimal volatility and variation in terms of energy related growth and 
##### development. Overall, said data is also based on current energy policy within
##### the United States. There are many more scenarios/cases but for the sake of 
##### wordcount and codelines, this report limits its analyzatin to only the described
##### reference case in order to explore how energy is set to change moving forward.


#### **US Consumption vs Production Forecasts**

##### The figure below demonstrates EIA Reference case forecasts on energy consumption
##### and production from 2020 to 2050. One observation from the data is that
##### renewable energy is set to grow by roughly 200% from 2020 to 2050. While that is
##### a positive observation as it relates to how energy consumption and production is
##### made up, it can be observed that both [natural gas] and [liquid fuels] are set to 
##### make up the majority of energy consumption up until 2050. 
```{r US Production/Consumption plots, echo=FALSE, include = TRUE}
# here we are plotting energy consumption and production and differentiating by
# if the energy data is consumption or production
# overall we don't need total values because it makes data less readable
usenergy %>%
  filter (fueltype != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                            y = quads, 
                            color = fueltype)) + 
  labs(x = "Year", y = "quads", title = "Energy Consumption by Fuel Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom") +
  facet_grid(cols=vars(result))

```



#### **US Export vs Import Forecasts**

##### Petroleum Exports are expected to peak around 2026 and then steadily decline by 
##### 2050. Natural Gas Exports on the other hand are expexcted to steadily rise while 
#### Coal exports are expected to decline very slowly to 2050. 

#### In terms of imports, Crude Oil is expected to, on average, increase in terms of imports by 2050. All other fuel types are expected to decline. 

#### Looking at export/import data forecasts for the US, it can be assumed that overall, 
#### US current energy policy will be successfull in leading to a decline of trade with 
#### non-renewable fuels, although the decline will be very slow. Overall, up until 2050,
#### trade of fossil fuels is expected to reman relatively high. 
```{r US Import/Export Plots, echo=FALSE}
# plotting the binded df
export_import %>%
  filter (fueltype != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                            y = quads, 
                            color = fueltype)) + 
  labs(x = "Year", y = "quads", title = "Energy Exports/Imports by Fuel Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom") +
  facet_grid(cols = vars(direction))

```


#### **US Energy Prices Forecasts**

##### Lookinga the the included figure, both BRENT CRUDE and WTI CRUDE are expected to
##### peak real prices by 2050. Said indexes are expected to reach the $90+ range. 
##### Referring back to the US energy production graph, this makes sense given there is
##### set to to be a very small decrease in Liquid Fuels Production moving forward. 
##### Moreover, one must acknowlege the global market when taking into consideration 
##### energy prices. It is expected that many nations will phase out fossil fuels 
##### moving forward. This can be expected to drive up oil prices moving forward. 
##### What is interesting is that electricity prices are set to slightly decrease 
##### moving forward until 2050.

```{r US Energy Prices Plot, echo=FALSE}
# plotting indexes forecast
energy_real_prices %>%
  ggplot() + geom_line(aes(x = year, 
                           y = price, 
                           color = index)) + 
  labs(x = "Year", y = "price", title = "Energy Real Prices by Index, 
       USA 2020 - 2050") + theme(legend.position = "bottom")

```


#### **US Emissions Forecasts**

##### Emissions related to Petroleum and Natural Gas are expected to rise in the US
##### up until 2050. On the other hand, coal is set to slowly decline from approximatley
##### 880 MMmt of CO2 to roughly below 650 MMmt of CO2 in 2050. This expected decline 
##### alongside the aforementioned Fuel Type emissions increases seem to indicate an
##### overall decrease of US emissions.
```{r US Emissions Plot, echo=FALSE}
# plotting emissions by fuel type
emissions %>%
  filter (fueltype != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                           y = mmotco2, 
                           color = fueltype)) + 
  labs(x = "Year", y = "MMmt CO2", title = "Carbon Emissions by Fuel Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom")

```


#### **US Renewable Energy Forecasts**

##### The data in this figure indicates that Solar and Wind energy technologies
##### are set to grow exponentially over the next few decades. For example, as it
##### relates to Solar PV technology, there is an expected 300 Gw increase from 
#### 2020 to 2050. That is roughly a %600 increase from 2020 levels.
```{r Renewable Energy Plots, echo=FALSE}
# binded plots differentiated by status
capacity_overview %>%
  filter (type != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                           y = gw, 
                           color = type)) + 
  labs(x = "Year", y = "Gigawatts", title = "Energy Capacity by Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom") +
  facet_grid(cols = vars(status))

```


#### **Global Energy and Emissions Forecast**

##### As expected, non-OECD countries are expected to grow at an alarming rate
##### in terms of how much energy is consumed. This can be assume to be partially
##### due to economic development and population growth. OECD energy use is expected
##### to remain under 300 quads Btu in 2050, whereas non-OECD energy use is expected
##### to reach above 600 quads Btu, essentially double the energy use of OECD countries.

##### It can be observed from the World Carbon Emissions graph below that non-OECD
##### Asian countries are expected to make up the majority of carbon emissions by 2050.
##### OECD countries are essentially projected to remain in slow decline by 2050. 
```{r Global plots, echo=FALSE}
# plotting global energy use
energy_use_global %>%
  ggplot() + geom_line(aes(x = year, 
                           y = quadbtu,
                           color = region)) + 
  labs(x = "Year", y = "quad Btu", title = "Global Energy Use, 2010 - 2050") +
  theme(legend.position = "bottom")

# plotting oecd emissions
world_emissions %>%
  filter(region != "total") %>%
  ggplot() + geom_line(aes(x = year,
                           y = MMmt, 
                           color = region)) + 
  labs(x = "Year", y = "MMmt CO2", title = "Global Carbon Emissions by Economic Region,
       2010 - 2050") + theme(legend.position = "bottom") +
  facet_grid(cols=vars(econ_type))

```

#### **Conclusion**

##### In conclusion, the above data presents a pretty unexciting trend of energy 
##### moving forward. Interesting developments that can be observed to be positive
##### include a rapid expansion of Renewable Energy Technology and Capacity.
##### As exciting as that is however, there is the fact that non-renewable fuels 
##### will remain very present in our energy mix up until 2050. Moreover, there is
##### an observed hurdle awaiting us in the future related to emsissions growth. 
##### It is forecasted that emissions from non-OECD countries is set to be double
##### that of OECD countries. This is bad news as it foreshadows a failure on OECD
##### countries to provide financial and technological support for emerging economies
##### related to energy use and production. Thus, overall, renewable technology is not
##### expected to growth at a fast enough pace given current energy policies in the US.
##### Lastly, energy use is expected to continue growing and at a consisten pace, meaning
##### that, given population growth and developing economy growth, the issue related
##### to energy consumption and emissions can be assumed to only continue to get bigger. 
