---
title: "US & Global Energy Forecast Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# importing necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
```


```{r data import, include = FALSE}
# total energy production data
tot_energy_production <- read.csv("data/Total_Energy_Production.csv")

# total energy use data
tot_energy_use <- read.csv("data/Total_Energy_Use.csv")

# total energy imports data
tot_energy_imports <- read.csv("data/Total_Energy_Imports.csv")

# total energy exports data
tot_energy_exports <- read.csv("data/Total_Energy_Exports.csv")

# total energy real prices data
tot_energy_real_prices <- read.csv("data/Total_Energy_Real_Prices.csv")

# CO2 emissions by Fuel type
emissions_by_fuel_type <- read.csv("data/Carbon_Dioxide_Total_by_Fuel.csv")

# Renewable Energy generating capacity
renewable_capacity <- read.csv("data/Renewable_Energy_Generating_Capacity.csv")

# renewable energy capacity additions
renewable_additions <- read.csv("data/Renewable_Energy_Capacity_Additions.csv")

# energy use oecd vs non-oecd
global_energy_use <- read.csv("data/energy_use_OECD_non_OECD.csv")

# emissions by OECD
emissions_from_OECD <- read.csv("data/Carbon_dioxide_emissions_OECD.csv")

# emissions non OECD
emissions_from_non_OECD <- read.csv("data/Carbon_dioxide_emissions_non_OECD.csv")
```

```{r data transformation, include = FALSE}
#transforming energy production data - LETS GOOOO
new_energy_production <- tot_energy_production %>%
  select(year = Year, 
         liquidfuels = Crude.Oil.and.Lease.Condensate.quads,
         naturalgas = Natural.Gas.Plant.Liquids.quads,
         coal = Coal.quads,
         nuclear = Nuclear.quads,
         hydropower = Hydropower.quads,
         biomass = Biomass.quads,
         other_renewable_energy = Other.Renewable.Energy.quads,
         other = Other.quads,
         total = Total.quads) %>%
  pivot_longer(c("liquidfuels", "naturalgas", "coal", "nuclear", "hydropower", 
                 "biomass", "other_renewable_energy", "other", "total"), 
               names_to = "fueltype", values_to = "quads") %>%
    drop_na() # removing any NA values to avoid error message


# Checking new df
new_energy_production

# transforming energy consumption data
new_energy_use <- tot_energy_use %>%
  select(year = Year, 
         liquidfuels = Liquid.Fuels.quads,
         naturalgas = Natural.Gas.quads,
         coal = Coal.quads,
         nuclear = Nuclear.quads,
         hydropower = Hydropower.quads,
         biomass = Biomass.quads,
         other_renewable_energy = Other.Renewable.Energy.quads,
         other = Other.quads,
         total = Total.quads) %>%
  pivot_longer(c("liquidfuels", "naturalgas", "coal", "nuclear", "hydropower", 
                 "biomass", "other_renewable_energy", "other", "total"), 
               names_to = "fueltype", 
               values_to = "quads") %>%
    drop_na()

# checking new df
new_energy_use

# transforming import data
new_energy_imports <- tot_energy_imports %>%
  select(year = Year, 
         crudeoil = Crude.Oil.quads,
         liquidfuels = Liquid.Fuels.quads,
         naturalgas = Natural.Gas.quads,
         other = Other.quads,
         total = Total.quads) %>%
  pivot_longer(c("crudeoil", "liquidfuels", "naturalgas", "other", "total"), 
               names_to = "fueltype", 
               values_to = "quads") %>%
    drop_na()

# check new imports df
new_energy_imports

# transforming export data
new_energy_exports <- tot_energy_exports %>%
  select(year = Year,
         petroleum = Petroleum.quads,
         naturalgas = Natural.Gas.quads,
         coal = Coal.quads,
         total = Total.quads) %>%
  pivot_longer(c("petroleum", "naturalgas", "coal", "total"), 
               names_to = "fueltype", 
               values_to = "quads") %>%
    drop_na()

# checking new exports df
new_energy_exports

# transforming real prices forecast
energy_real_prices <- tot_energy_real_prices %>%
  select(year = Year, 
         brent = Brent..2020...b..2020...b,
         wti = West.Texas.Intermediate..2020...b..2020...b,
         gas_henry_hub = Gas.Price.at.Henry.Hub..2020...MMBtu..2020...MMBtu,
         coal_minemouth_ton = Coal.at.Minemouth..2020...ton..2020...ton,
         coal_minemouth = Coal.Minemouth..2020...MMBtu..2020...MMBtu,
         deliver_coal = Delivered.Coal..2020...MMBtu..2020...MMBtu,
         electricity = Electricity..2020.cents.kWh..2020.cents.kWh) %>%
  pivot_longer(c("brent", "wti", "gas_henry_hub", "coal_minemouth", 
                 "coal_minemouth_ton", "deliver_coal","electricity"),
               names_to = "index", 
               values_to = "price") %>%
    drop_na()

# checking energy real prices df
energy_real_prices

# transform emissions data
emissions <- emissions_by_fuel_type %>%
  select(year = Year, 
         petroleum = Petroleum.MMmt.CO2,
         naturalgas = Natural.Gas.MMmt.CO2,
         coal = Coal.MMmt.CO2,
         other = Other.MMmt.CO2,
         total = Total.MMmt.CO2) %>%
  pivot_longer(c("petroleum", "naturalgas", "coal", "other", "total"), 
               names_to = "fueltype", 
               values_to = "mmotco2") %>%
    drop_na()

# checking emissions df
emissions

# transform renewable energy capacity
capacity <- renewable_capacity %>%
  select(year = Year, 
         hydropower = Hydropower.GW,
         geothermal = Geothermal.GW,
         mun_waste = Municipal.Waste.GW, 
         biomass = Wood.and.Other.Biomass.GW,
         solar_thermal = Solar.Thermal.GW,
         solar_pv = Solar.Photovoltaic.GW,
         wind = Wind.GW,
         offshore = Offshore.Wind.GW,
         total = Total.GW) %>%
  pivot_longer(c("hydropower", "geothermal", "mun_waste", "biomass", 
                 "solar_thermal", "solar_pv", "wind","offshore", "total"),
               names_to = "type", 
               values_to = "gw") %>%
    drop_na()

# check capacity df
capacity

# transform capacity additions 
cap_adds <- renewable_additions %>%
  select(year = Year, 
         hydropower = Hydropower.GW,
         geothermal = Geothermal.GW,
         mun_waste = Municipal.Waste.GW, 
         biomass = Wood.and.Other.Biomass.GW,
         solar_thermal = Solar.Thermal.GW,
         solar_pv = Solar.Photovoltaic.GW,
         wind = Wind.GW,
         offshore = Offshore.Wind.GW,
         total = Total.GW) %>%
  pivot_longer(c("hydropower", "geothermal", "mun_waste", "biomass", 
                 "solar_thermal", "solar_pv", "wind","offshore", "total"),
               names_to = "type", 
               values_to = "gw")%>%
    drop_na()

# check capacity additions df
cap_adds

# transform data for global energy use
energy_use_global <- global_energy_use %>%
  select(year = Year, 
         oecd_total = OECD..Total.OECD.quad.Btu,
         non_oecd_total = Non.OECD..Total.Non.OECD.quad.Btu) %>%
  pivot_longer(c("oecd_total", "non_oecd_total"),
               names_to = "region", 
               values_to = "quadbtu") %>%
    drop_na()

# check energy global use data frame
energy_use_global

# transform oecd emissions
oecd_emissions <- emissions_from_OECD %>%
  select(year = Year, 
         americas = OECD.Americas.MMmt,
         europe = OECD.Europe.MMmt,
         asia = OECD.Asia.MMmt,
         total = Total.OECD.MMmt) %>%
  pivot_longer(c("americas", "europe", "asia", "total"),
               names_to = "region",
               values_to = "MMmt") %>%
    drop_na()

# check oecd emissions df
oecd_emissions

# transform non oecd emissions
non_oecd_emissions <- emissions_from_non_OECD %>%
  select(year = Year, 
         non_eur = Non.OECD.Europe.and.Eurasia.MMmt,
         non_asia = Non.OECD.Asia.MMmt,
         middleeast = Middle.East.MMmt,
         africa = Africa.MMmt,
         non_americas = Non.OECD.Americas.MMmt,
         total = Total.Non.OECD.MMmt) %>%
  pivot_longer(c("non_eur", "non_asia", "middleeast", "africa", "non_americas",
                 "total"),
               names_to = "region",
               values_to = "MMmt")%>%
    drop_na()

# check oecd emissions df
non_oecd_emissions

```

```{r deep data transformation, include = FALSE}
# here, we want to put import/export data together in one df
exports <- new_energy_exports %>%
  mutate(direction = "export")

# show data frame
exports

# here we are doing the same but for import
imports <- new_energy_imports %>%
  mutate(direction = "import")

# show imports data frame now
imports

# joining the two data sets
export_import <- rbind(exports, imports)

# show joined data frame 
export_import

```

#### **Energy**
##### The purpose of this report is to use publicly available data from the Energy
##### Information Administration (EIA) in order to think about the following questions:
1. Within a U.S. context, what do we expect energy to look like up until 2050?
2. Within a U.S. context, where is Renewable Energy headed?
3. Within a global context, what will energy and emissions look like thinking 
forward to 2050?


```{r US Production/Consumption plots, echo=FALSE, include = TRUE}
# fix the data of total energy production
# excluding total type given it only makes data less presentable
# plotting production
new_energy_production %>%
  filter (fueltype != "total") %>%
  ggplot(xmin = 2020, xmax = 2050) + geom_line(aes(x = year, 
                            y = quads, 
                            color = fueltype)) + 
  labs(x = "Year", y = "quads", title = "Energy Production by Fuel Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom")

# plotting energy consumption
new_energy_use %>%
  filter (fueltype != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                            y = quads, 
                            color = fueltype)) + 
  labs(x = "Year", y = "quads", title = "Energy Consumption by Fuel Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom")

```

```{r US Import/Export Plots, echo=FALSE}
# plotting imports
new_energy_imports %>%
  filter (fueltype != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                            y = quads, 
                            color = fueltype)) + 
  labs(x = "Year", y = "quads", title = "Energy Imports by Fuel Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom")

# plotting exports
new_energy_exports %>%
  filter (fueltype != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                            y = quads, 
                            color = fueltype)) + 
  labs(x = "Year", y = "quads", title = "Energy Exports by Fuel Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom")

# plotting the binded df
export_import %>%
  filter (fueltype != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                            y = quads, 
                            color = fueltype)) + 
  labs(x = "Year", y = "quads", title = "Energy Exports/Imports by Fuel Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom") +
  facet_grid(cols = vars(direction))

```

```{r US Energy Prices Plot, echo=FALSE}
# plotting indexes forecast
energy_real_prices %>%
  ggplot() + geom_line(aes(x = year, 
                           y = price, 
                           color = index)) + 
  labs(x = "Year", y = "price", title = "Energy Real Prices by Index, 
       USA 2020 - 2050") + theme(legend.position = "bottom")

```

```{r US Emissions Plot, echo=FALSE}
# plotting emissions by fuel type
emissions %>%
  filter (fueltype != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                           y = mmotco2, 
                           color = fueltype)) + 
  labs(x = "Year", y = "MMmt CO2", title = "Carbon Emissions by Fuel Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom")

```


```{r Renewable Energy Plots, echo=FALSE}
# plotting capacity by type
capacity %>%
  filter (type != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                           y = gw, 
                           color = type)) + 
  labs(x = "Year", y = "Gigawatts", title = "Energy Capacity by Type, 
       USA 2020 - 2050") + theme(legend.position = "bottom")

# plotting capacity additions  by type
cap_adds %>%
  filter (type != "total") %>%
  ggplot() + geom_line(aes(x = year, 
                           y = gw, 
                           color = type)) + 
  labs(x = "Year", y = "Gigawatts", title = "Energy Capacity Additions by Type, 
       USA 2021 - 2050") + theme(legend.position = "bottom")

```


```{r Global plots, echo=FALSE}
# plotting global energy use
energy_use_global %>%
  ggplot() + geom_line(aes(x = year, 
                           y = quadbtu,
                           color = region)) + 
  labs(x = "Year", y = "quad Btu", title = "Global Energy Use, 2010 - 2050") +
  theme(legend.position = "bottom")

# plotting oecd emissions
oecd_emissions %>%
  filter(region != "total") %>%
  ggplot() + geom_line(aes(x = year,
                           y = MMmt, 
                           color = region)) + 
  labs(x = "Year", y = "MMmt CO2", title = "OECD Carbon Emissions by Region,
       2010 - 2050") + theme(legend.position = "bottom")

# plotting non - oecd emissions
non_oecd_emissions %>%
  filter(region != "total") %>%
  ggplot() + geom_line(aes(x = year,
                           y = MMmt, 
                           color = region)) + 
  labs(x = "Year", y = "MMmt CO2", title = "Non-OECD Carbon Emissions by Region,
       2010 - 2050") + theme(legend.position = "bottom")

```

