---
title: "Exploring U.S. Supreme Court Decisions"
author: "Angel Salazar"
output: github_document
---

## Get the data

```{r}
# load useful packages
library(tidyverse)

# load data
scdb_case <- read.csv("data/scdb-case.csv")
scdb_vote <- read.csv("data/scdb-vote.csv")
```

## Notes 
1. 'scdb-case.csv' contains all *case-level* variables
    - contains one row for every case and one column for every variable
2. 'scdb-vote.csv' contains all *vote-level* variables
    - contains one row for every vote by a justice in every case and one
        column for every variable
3. For our purposes, the only ID variable we need to concern ourselves with is *caseIssuesId* but these may also be useful:
    - chief
    - dateDecision
    - decisionDirection
    - decisionType
    - declarationUncon
    - direction
    - issueArea
    - justice
    - justiceName
    - majority
    - majVotes
    - minVotes
    - term

## Recode variables as you find necessary

```{r}
# left join df
scdb_left_join <- left_join(x = scdb_case, 
                            y = scdb_vote, 
                            by = NULL)

# right join df
scdb_right_join <- right_join(x = scdb_case, 
                              y = scdb_vote, 
                              by = NULL)

# inner join df
scdb_inner_join <- inner_join(x = scdb_case, 
                              y = scdb_vote, 
                              by = NULL)

# full join df
scdb_full_join <- full_join(x = scdb_case,
                            y = scdb_vote, 
                            by = NULL)
```

## What percentage of cases in each term are decided by a one-vote margin (i.e. 5-4, 4-3, etc.)

```{r}
# Here we will essentially have to filter out any data where each term is not
# decided by a one-vote margin. From the included variables to consider in our
# notes, the most relevant are (1) majority, (2) majVotes, and (3) minVotes
# best way to calculate one-vote margin is => majVotes - minVotes
# something to take into account is that term runs from 1791 to 2019

# setting up a data framework to hold only the one-vote margins
one_vote_margin <- filter(scdb_case, 
                                    majVotes - minVotes == "1")

# we also need to consider each term

scdb_case %>%
  
```

## For justices [currently serving on the Supreme Court](https://www.supremecourt.gov/about/biographies.aspx), how often have they voted in the conservative direction in cases involving criminal procedure, civil rights, economic activity, and federal taxation?

Organize the resulting graph by justice in descending order of seniority. (Note that the chief justice is always considered the most senior member of the court, regardless of appointment date.)

```{r}

```

## In each term, how many of the term's published decisions (decided after oral arguments) were announced in a given month?

```{r}

```

## Which justices are most likely to agree with with the Court's declaration that an act of Congress, a state or territorial law, or a municipal ordinance is unconstitutional? Identify all cases where the Court declared something unconstitutional and determine the ten justices who most and least frequently agreed with this outcome as a percentage of all votes cast by the justice in these cases. Exclude any justice with fewer than 30 votes in cases where the Court's outcome declares something unconstitutional.

```{r}

```

## In each term he served on the Court, in what percentage of cases was Justice Antonin Scalia in the majority?

```{r}

```

## Create a graph similar to above that adds a second component which compares the percentage for all cases versus non-unanimous cases (i.e. there was at least one dissenting vote)

```{r}

```

## In each term, what percentage of cases were decided in the conservative direction?

```{r}

```

## The Chief Justice is frequently seen as capable of influencing the ideological direction of the Court. Create a graph similar to the one above that also incorporates information on who was the Chief Justice during the term.

```{r}

```

## Session info

```{r, echo = TRUE}
devtools::session_info()
```


